# manuscript-code
Code for the manuscript: High-throughput drug screening identifies BCL-xL targeting as a potent strategy for radiosensitization in rhabdomyosarcoma tumoroids 

---
title: "RMS_EBRT_Github"
author: "Se Whee Sammy Park"
date: "2025-04-29"
output: html_document
---

High-throughput drug screening identifies BCL-xL targeting as a potent strategy for radiosensitization in rhabdomyosarcoma tumoroids: Main Figures

# Load Packages
Load all required package
```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(ggpubr)
library(ggrepel)
library(scales)
library(RColorBrewer)
library(data.table)
library(reshape2)
library(plyr)
library(rstatix)
library(ComplexHeatmap)
library(circlize)
library(drc)
library(readxl)
library(readr)
library(tidyr)
library(stringr)
library(grid)
library(gridExtra)
library(ggpattern)
library(synergyfinder)
```

# Figure 2
## Figure 2A
Scatterplot of DSS between Karolinska Institue vs Princess Maxima Centrum
```{r, FIG 2A}
setwd()

# Choose and merge data sets
df_KI <- read.table("FIG_2A_RMS_DrugScreen_KI.csv",sep=",", header = TRUE)
df_PMC <- read.table("FIG_2A_RMS_DrugScreen_PMC.csv",sep=",", header = TRUE)
df_merge <- merge(df_KI,df_PMC,by="Drug_Name")

# Plot a scatter plot - RMS006
ggscatter(df_merge, x = "RMS006", y = "RMS006_KI",
          add = "reg.line",  # Add regression line
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          conf.int = TRUE ) + # Add confidence interval
  stat_cor(method = "spearman", label.x = 1, label.y = 40) +
  geom_text_repel(data = subset(df_merge,RMS006_KI > 30), 
                  aes(label = paste0(Drug_Name)) ,
                  hjust = 2,
                  nudge_x = 2,
                  direction = "y",
                  fontface = "bold",
                  size = 4) +
  geom_text_repel(data = subset(df_merge, RMS006 > 30 & RMS006_KI < 30),
                  aes(label = paste0(Drug_Name)) ,
                  hjust = 2,
                  nudge_x = 2,
                  direction = "y",
                  fontface = "bold",
                  size = 4) +
  labs(x="RMS006_EARLY", y="RMS006_LATE") 

# Save plot - RMS006
ggsave("FIG_2A_Spearman_correlation_RMS006_4_4.pdf", width = 4, height = 4, dpi = 600)

# Plot a scatter plot - RMS102
ggscatter(df_merge, x = "RMS102", y = "RMS102_KI",
          add = "reg.line",  # Add regression line
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          conf.int = TRUE ) + # Add confidence interval
  stat_cor(method = "spearman", label.x = 1, label.y = 40) +
  geom_text_repel(data = subset(df_merge,RMS102_SP > 30), 
                  aes(label = paste0(Drug_Name)) ,
                  hjust = 2,
                  nudge_x = 2,
                  direction = "y",
                  fontface = "bold",
                  size = 4) +
  geom_text_repel(data = subset(df_merge, RMS102 > 30 & RMS102_SP < 30),
                  aes(label = paste0(Drug_Name)) ,
                  hjust = 2,
                  nudge_x = 2,
                  direction = "y",
                  fontface = "bold",
                  size = 4) +
  labs(x="RMS102_EARLY", y="RMS102_LATE") 

# Save plot - RMS102
ggsave("FIG_2A_Spearman_correlation_RMS102_4_4.pdf", width = 4, height = 4, dpi = 600)

# Plot a scatterplot - RMS410
ggscatter(df_merge, x = "RMS410", y = "RMS410_KI",
          add = "reg.line",  # Add regression line
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          conf.int = TRUE ) + # Add confidence interval
  stat_cor(method = "spearman", label.x = 1, label.y = 40) +
  geom_text_repel(data = subset(df_merge,RMS410_SP > 30), 
                  aes(label = paste0(Drug_Name)) ,
                  hjust = 2,
                  nudge_x = 2,
                  direction = "y",
                  fontface = "bold",
                  size = 4) +
  geom_text_repel(data = subset(df_merge, RMS410 > 30 & RMS410_SP < 30),
                  aes(label = paste0(Drug_Name)) ,
                  hjust = 2,
                  nudge_x = 2,
                  direction = "y",
                  fontface = "bold",
                  size = 4) +
  labs(x="RMS410_EARLY", y="RMS410_LATE") 

# Save plot - RMS410
ggsave("FIG_2A_Spearman_correlation_RMS410_4_4.pdf", width = 4, height = 4, dpi = 600)
```

## Figure 2B
Drug DSS comparison between RMS tumoroids and BM
```{r, FIG 2B}
setwd()

# Choose data 
df_KI  <- read.table("FIG_2A_RMS_DrugScreen_KI.csv",sep=",", header = TRUE)

# Remove outlier drug - Everolimus
df_KI <- subset(df_KI,Drug_Name != "Everolimus") #Everolimus drug curves not reliable

# Reannotate the data frame row and columns
rownames(df_KI) <- df_KI[,1]
df_KI_small <- df_KI[1:20,]
df_KI_m <- as.matrix(df_KI[, 6:15])
df_KI_m_small <- df_KI_m[1:20,]
anno = anno_mark(at = c(1:20), labels = df_KI_small$Mechanism, which = "row")
column_ha = HeatmapAnnotation(foo = c(0,0,0,0,0,1,1,1,1,1) )

# Plot a heatmap
pdf("FIG_2B_ComplexHeatmap_Top20_KI_drugs_noEvero.pdf", width = 8.5, height = 8)
ht <- Heatmap(df_KI_m_small,
              cluster_rows=FALSE,
              cluster_columns=FALSE,
              top_annotation = column_ha,
              col=colorRamp2(c(0, 15, 30),c("blue", "white", "red")),
              heatmap_legend_param = list(
                legend_direction = "horizontal", 
                legend_width = unit(5, "cm")
              )
)+
  rowAnnotation(mark = anno)

draw(ht, auto_adjust = FALSE,heatmap_legend_side = "bottom")
dev.off()
```

## Figure 2C 
Drug target DSS comparison between RMS tumoroids and BM
```{r, FIG 2C}
setwd()

# Choose data
df <- read.table("FIG_2C_RMS_drugScreen_geneCentric.txt", header = TRUE)

# Reannotate df
head(df)
df$FIMM_ID <- NULL
df$Class_Explained <- NULL
df$Clinical_Phase <- NULL
df$Ave.RMS <- NULL
df$Ave.BM <- NULL
df$Diff <- NULL
df$RMS.SD <- NULL
df2 <- df

# Wilcoxon signed-rank test
df2$wilcox_p <- apply(df2[, 4:13], 1, function(x) wilcox.test(x[1:5], x[6:10])$p.value)
write.csv(df2, "RMS_drugScreen_geneCentric_wilcoxon.csv")

# Rearrage and reannotate dataframe
long <- melt(df, id.vars = c("Drug_Name", "Gene_Targeted", "Mechanism"))
long$samples <- long$variable
long$samples <- as.character(long$samples)
long$samples <- removeNumbers(long$samples)
long$Gene_Targeted <- factor(long$Gene_Targeted, levels=c('PLK1', 'BCL2L1', 'MDM2', 'MTOR',"KIF11",
                                                          "CDH16","HMGCR","TP53","PIK3CA","PIK3CG",
                                                          "IGF1R","AURKA","FKBP1A","DHODH","FNTA",
                                                          "AURKC","ESR1","ESR2","AKT1","PRKACA"))
long$Gene_Targeted <- fct_rev(long$Gene_Targeted)
long  <- na.omit(long)
my_comparisons <- list( c("RMS", "BM.") )

# Plot a boxplot 
ggplot(long, aes(x=Gene_Targeted, y=value, fill = samples)) + 
  geom_boxplot(outlier.shape = NA, position=position_dodge(1)) +
  scale_fill_brewer(palette="Dark2")+
  geom_point(color="black",size=0.6, alpha=0.9, position=position_jitterdodge(dodge.width=1.0,
                                                                              jitter.width = 0.1)) +
  stat_compare_means(method = "wilcox.test", paired=FALSE, label.x = 35, label = "p.signif") +
  
  theme(axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position="bottom",
        panel.background = element_rect(fill = 'white')) +
  coord_flip()

# Save plot
ggsave(filename = "FIG_2C_GeneCentric_Boxplot_Top20.pdf", width = 3, height = 4, units = "in", device='pdf')
```

# Figure 3
## Figure 3A
Scatterplot for DSS with and without irradiation - RMS006
```{r, FIG 3A}
setwd()

# Choose data
df <- read.table("FIG_3A_RMS_EBRT_screen_and_BM.csv",sep=",", header = TRUE)

# Plot a scatterplot - RMS006
ggscatter(df, x = "RMS006", y = "RMS006_EBRT",
          add = "reg.line",  # Add regression line
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          conf.int = TRUE ) + # Add confidence interval
  stat_cor(method = "spearman", label.x = 1, label.y = 40) +
  geom_text_repel(data = subset(df,RMS006_EBRT-RMS006 > 15), 
                  aes(label = paste0(Drug.Name)) ,
                  hjust = 2,
                  nudge_x = 2,
                  direction = "y",
                  fontface = "bold",
                  size = 4) +
  geom_text_repel(data = subset(df, RMS006-RMS006_EBRT > 10),
                  aes(label = paste0(Drug.Name)) ,
                  hjust = -2,
                  nudge_x = 2,
                  direction = "y",
                  fontface = "bold",
                  size = 4) +
  geom_text_repel(data = subset(df, RMS006 > 40),
                  aes(label = paste0(Drug.Name)) ,
                  hjust = -2,
                  nudge_x = 2,
                  direction = "y",
                  fontface = "bold",
                  size = 4) +
  labs(x="RMS006", y="RMS006_EBRT") 

# Save plot - RMS006
ggsave(filename = paste0("FIG_3A_RMS006_X_Y_4_4.pdf"), width = 4, height = 4, units = "in", device='pdf')
```

## Figure 3B
Scatterplot for DSS with and without irradiation - RMS410
```{r, FIG 3B}
setwd()

# Choose data
df <- read.table("FIG_3A_RMS_EBRT_screen_and_BM.csv",sep=",", header = TRUE)

# Plot a scatterplot - RMS410
ggscatter(df, x = "RMS410", y = "RMS410_EBRT",
          add = "reg.line",  # Add regression line
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          conf.int = TRUE ) + # Add confidence interval
  stat_cor(method = "spearman", label.x = 1, label.y = 40) +
  geom_text_repel(data = subset(df,RMS410_EBRT-RMS410 > 7), 
                  aes(label = paste0(Drug.Name)) ,
                  hjust = 2,
                  nudge_x = 2,
                  direction = "y",
                  fontface = "bold",
                  size = 4) +
  geom_text_repel(data = subset(df, RMS410-RMS410_EBRT > 7),
                  aes(label = paste0(Drug.Name)) ,
                  hjust = -2,
                  nudge_x = 2,
                  direction = "y",
                  fontface = "bold",
                  size = 4) +
  geom_text_repel(data = subset(df, RMS410 > 35),
                  aes(label = paste0(Drug.Name)) ,
                  hjust = -2,
                  nudge_x = 2,
                  direction = "y",
                  fontface = "bold",
                  size = 4) +
  labs(x="RMS410", y="RMS410_EBRT") 

# Save plot - RMS410
ggsave(filename = paste0("FIG_3A_RMS410_X_Y_7_7_4_4.pdf"), width = 4, height = 4, units = "in", device='pdf')
```

## Figure 3C
Top radiosensitizers and radioprotectors
```{r, FIG 3C}
setwd()

# Choose data
df <- read.table("FIG_3C_RMS_EBRT_screen_dDSS.csv",sep=",", header = TRUE)

# Reannotate the dataframe
df$FIMM.ID <- NULL
df$Clinical.Phase <- NULL
df$Class.Explained <- NULL
df$Clinical.Phase <- NULL

# Define the top ten radiosensitizing and radioprotecting compounds
df_top <- df[order(-df$average),] 
df_top10 <- df_top[1:10,]
df_bottom <- df[order(df$average),] 
df_bottom10 <- df_bottom[1:10,]

# Merge the top hits
df_merge <- rbind(df_top10,df_bottom10)

# Order data by drug name
rownames(df_order) <- df_order$Drug.Name

# Rearrange the data
df_merge_long <- melt(setDT(df_merge), id.vars = c("Drug.Name","Mechanism","average"), variable.name = "Viability")

# Calculate the mean and SD of each Drug
df_merge_long_summary <- ddply(df_merge_long, c("Drug.Name","Mechanism","average"), summarise,
                               mean = mean(value), sd = sd(value))
df_merge_long_summary$mean <- as.numeric(df_merge_long_summary$mean)
df_merge_long_summary$sd <- as.numeric(df_merge_long_summary$sd)
df_merge_long_summary$Drug.Name <- reorder(df_merge_long_summary$Drug.Name, df_merge_long_summary$average)
df_merge_long_summary <- df_merge_long_summary[order(-df_merge_long_summary$average),] 

# Plot barplot
ggplot(df_merge_long_summary, aes(x=average,y=Drug.Name, fill=Mechanism))+
  geom_bar(stat="identity") +
  geom_errorbar(aes(xmin=mean-sd, xmax=mean+sd), width=.2,
                position=position_dodge(0.05))+
  theme(axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=0),
        panel.background = element_rect(fill = 'white'))+
  coord_flip()

# Save plot
ggsave(filename = "FIG_3C_RMS_EBRT_dDSS_barchart_topBottom10_combined.pdf", width = 8, height = 8, units = "in", device='pdf')
```

## Figure 3D
Average dose response curves - BCL-xL inhibitors
```{r, FIG 3D}
setwd()

# Choose drug
drug <-"WEHI-539"
drug <-"A-1331852"
drug <-"Navitoclax"
drug <-"A-1155463"

# Choose data referring to the drug of choice and annotate model, irradiation status and EC50
df_006 <- read.table("FIG_3D_FO5A_RMS006_DSRT.csv",sep=",", header = TRUE)
df_006_EBRT <- read.table("FIG_3D_FO5A_EBRT_RMS006_DSRT.csv",sep=",", header = TRUE)
df_006_drug <- subset(df_006, DRUG_NAME == drug)
df_006_EBRT_drug <- subset(df_006_EBRT, DRUG_NAME == drug)
df_006_drug$EBRT <- "no"
df_006_EBRT_drug$EBRT <- "yes"
df_006_drug$RMS <- "006"
df_006_EBRT_drug$RMS <- "006"
EC50_006 <- df_006_drug$EC50
EC50_006_EBRT <- df_006_EBRT_drug$EC50

# Choose data referring to the drug of choice and annotate model, irradiation status and EC50
df_102 <- read.table("FIG_3D_FO5A_RMS102_DSRT.csv",sep=",", header = TRUE)
df_102_EBRT <- read.table("FIG_3D_FO5A_EBRT_RMS102_DSRT.csv",sep=",", header = TRUE)
df_102_drug <- subset(df_102, DRUG_NAME == drug)
df_102_EBRT_drug <- subset(df_102_EBRT, DRUG_NAME == drug)
df_102_drug$EBRT <- "no"
df_102_EBRT_drug$EBRT <- "yes"
df_102_drug$RMS <- "102"
df_102_EBRT_drug$RMS <- "102"
EC50_102 <- df_102_drug$EC50
EC50_102_EBRT <- df_102_EBRT_drug$EC50

# Choose data referring to the drug of choice and annotate model, irradiation status and EC50
df_410 <- read.table("FIG_3D_FO5A_RMS410_DSRT.csv",sep=",", header = TRUE)
df_410_EBRT <- read.table("FIG_3D_FO5A_EBRT_RMS410_DSRT.csv",sep=",", header = TRUE)
df_410_drug <- subset(df_410, DRUG_NAME == drug)
df_410_EBRT_drug <- subset(df_410_EBRT, DRUG_NAME == drug)
df_410_drug$EBRT <- "no"
df_410_EBRT_drug$EBRT <- "yes"
df_410_drug$RMS <- "410"
df_410_EBRT_drug$RMS <- "410"
EC50_410 <- df_410_drug$EC50
EC50_410_EBRT <- df_410_EBRT_drug$EC50

# Choose a reference dataframe
df_drug <- df_006_drug

# Combine data
comb <- rbind(df_006_drug,df_102_drug,df_410_drug,df_006_EBRT_drug,df_102_EBRT_drug,df_410_EBRT_drug)

# Rearrange dataframe
long_comb <- melt(setDT(comb), id.vars = c("ID","DRUG_NAME","ANALYSIS_NAME","EC50","SLOPE","MAX","MIN",
                                           "Min.Conc.tested","Max.Conc.tested","EC50_std_error","EBRT","RMS"), variable.name = "Viability")

# Add the drug concentration
long_comb$concentration <- c(df_drug$Min.Conc.tested*1,df_drug$Min.Conc.tested*1,df_drug$Min.Conc.tested*1,df_drug$Min.Conc.tested*1,df_drug$Min.Conc.tested*1,df_drug$Min.Conc.tested*1,
                             df_drug$Min.Conc.tested*10,df_drug$Min.Conc.tested*10,df_drug$Min.Conc.tested*10,df_drug$Min.Conc.tested*10,df_drug$Min.Conc.tested*10,df_drug$Min.Conc.tested*10,
                             df_drug$Min.Conc.tested*100,df_drug$Min.Conc.tested*100,df_drug$Min.Conc.tested*100,df_drug$Min.Conc.tested*100,df_drug$Min.Conc.tested*100,df_drug$Min.Conc.tested*100,
                             df_drug$Min.Conc.tested*1000,df_drug$Min.Conc.tested*1000,df_drug$Min.Conc.tested*1000,df_drug$Min.Conc.tested*1000,df_drug$Min.Conc.tested*1000,df_drug$Min.Conc.tested*1000,
                             df_drug$Min.Conc.tested*10000,df_drug$Min.Conc.tested*10000,df_drug$Min.Conc.tested*10000,df_drug$Min.Conc.tested*10000,df_drug$Min.Conc.tested*10000,df_drug$Min.Conc.tested*10000)

# Calculate the mean and SD for each drug
long_comb_summary <- ddply(long_comb, c("concentration","EBRT", "Viability"), summarise,
                           mean = mean(value), sd = sd(value))

# Plot a dose response curve
ggplot(data = long_comb_summary, aes(x = concentration, y = mean, group=EBRT, color = EBRT)) + 
  geom_point(size = 3) +
  geom_line() +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                position=position_dodge(0.05))+
  scale_x_log10(labels = scales::comma) +
  scale_color_manual(values=c("yes"="blue","no"="red"))+
  ylim(-30, 130) +
  
  theme(axis.line = element_line(colour = "black"),legend.position="none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background = element_rect(fill = 'white'))


# Save the plot
ggsave(filename = paste0("FIG_3D_Drug_curve_EBRT_",drug,"_three_combined.pdf"), width = 3, height = 3, units = "in", device='pdf')
```


## Figure 3E
Average dose response curves - PLK1 inhibitors 
```{r, FIG 3E}
setwd()

# Choose drug
drug <-"GSK-461364"
drug <-"BI 2536"
drug <-"Volasertib"
drug <-"Methotrexate"

# Choose data referring to the drug of choice and annotate model, irradiation status and EC50
df_006 <- read.table("FIG_3D_FO5A_RMS006_DSRT.csv",sep=",", header = TRUE)
df_006_EBRT <- read.table("FIG_3D_FO5A_EBRT_RMS006_DSRT.csv",sep=",", header = TRUE)
df_006_drug <- subset(df_006, DRUG_NAME == drug)
df_006_EBRT_drug <- subset(df_006_EBRT, DRUG_NAME == drug)
df_006_drug$EBRT <- "no"
df_006_EBRT_drug$EBRT <- "yes"
df_006_drug$RMS <- "006"
df_006_EBRT_drug$RMS <- "006"
EC50_006 <- df_006_drug$EC50
EC50_006_EBRT <- df_006_EBRT_drug$EC50

# Choose data referring to the drug of choice and annotate model, irradiation status and EC50
df_102 <- read.table("FIG_3D_FO5A_RMS102_DSRT.csv",sep=",", header = TRUE)
df_102_EBRT <- read.table("FIG_3D_FO5A_EBRT_RMS102_DSRT.csv",sep=",", header = TRUE)
df_102_drug <- subset(df_102, DRUG_NAME == drug)
df_102_EBRT_drug <- subset(df_102_EBRT, DRUG_NAME == drug)
df_102_drug$EBRT <- "no"
df_102_EBRT_drug$EBRT <- "yes"
df_102_drug$RMS <- "102"
df_102_EBRT_drug$RMS <- "102"
EC50_102 <- df_102_drug$EC50
EC50_102_EBRT <- df_102_EBRT_drug$EC50

# Choose data referring to the drug of choice and annotate model, irradiation status and EC50
df_410 <- read.table("FIG_3D_FO5A_RMS410_DSRT.csv",sep=",", header = TRUE)
df_410_EBRT <- read.table("FIG_3D_FO5A_EBRT_RMS410_DSRT.csv",sep=",", header = TRUE)
df_410_drug <- subset(df_410, DRUG_NAME == drug)
df_410_EBRT_drug <- subset(df_410_EBRT, DRUG_NAME == drug)
df_410_drug$EBRT <- "no"
df_410_EBRT_drug$EBRT <- "yes"
df_410_drug$RMS <- "410"
df_410_EBRT_drug$RMS <- "410"
EC50_410 <- df_410_drug$EC50
EC50_410_EBRT <- df_410_EBRT_drug$EC50

# Choose a reference dataframe
df_drug <- df_006_drug

# Combine data
comb <- rbind(df_006_drug,df_102_drug,df_410_drug,df_006_EBRT_drug,df_102_EBRT_drug,df_410_EBRT_drug)

# Rearrange dataframe
long_comb <- melt(setDT(comb), id.vars = c("ID","DRUG_NAME","ANALYSIS_NAME","EC50","SLOPE","MAX","MIN",
                                           "Min.Conc.tested","Max.Conc.tested","EC50_std_error","EBRT","RMS"), variable.name = "Viability")

# Add the drug concentration
long_comb$concentration <- c(df_drug$Min.Conc.tested*1,df_drug$Min.Conc.tested*1,df_drug$Min.Conc.tested*1,df_drug$Min.Conc.tested*1,df_drug$Min.Conc.tested*1,df_drug$Min.Conc.tested*1,
                             df_drug$Min.Conc.tested*10,df_drug$Min.Conc.tested*10,df_drug$Min.Conc.tested*10,df_drug$Min.Conc.tested*10,df_drug$Min.Conc.tested*10,df_drug$Min.Conc.tested*10,
                             df_drug$Min.Conc.tested*100,df_drug$Min.Conc.tested*100,df_drug$Min.Conc.tested*100,df_drug$Min.Conc.tested*100,df_drug$Min.Conc.tested*100,df_drug$Min.Conc.tested*100,
                             df_drug$Min.Conc.tested*1000,df_drug$Min.Conc.tested*1000,df_drug$Min.Conc.tested*1000,df_drug$Min.Conc.tested*1000,df_drug$Min.Conc.tested*1000,df_drug$Min.Conc.tested*1000,
                             df_drug$Min.Conc.tested*10000,df_drug$Min.Conc.tested*10000,df_drug$Min.Conc.tested*10000,df_drug$Min.Conc.tested*10000,df_drug$Min.Conc.tested*10000,df_drug$Min.Conc.tested*10000)

# Calculate the mean and SD of each Drug
long_comb_summary <- ddply(long_comb, c("concentration","EBRT", "Viability"), summarise,
                           mean = mean(value), sd = sd(value))
# Plot the dose reponse curve
ggplot(data = long_comb_summary, aes(x = concentration, y = mean, group=EBRT, color = EBRT)) + 
  geom_point(size = 3) +
  geom_line() +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                position=position_dodge(0.05))+
  scale_x_log10(labels = scales::comma) +
  scale_color_manual(values=c("yes"="blue","no"="red"))+
  ylim(-30, 130) +
  
  theme(axis.line = element_line(colour = "black"),legend.position="none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background = element_rect(fill = 'white'))

# Save the plot
ggsave(filename = paste0("FIG_3E_Drug_curve_EBRT_",drug,"_three_combined.pdf"), width = 3, height = 3, units = "in", device='pdf')
```


# Figure 4
## Figure 4B
Heatmap of the validation screen
```{r, FIG 4B}
setwd()

# Choose data
df <- read.table("FIG_4B_Validation_dDSS_and_timing.csv",sep=",", header = TRUE)

# Define drug names as rows
rownames(df) <- df$DRUG_NAME

# Subset necessary dataframe
df_m <- as.matrix(df[, 4:16])

# Allocate colors to each sample
colors = list(bar = c("HEK293" = "#009ACD","HEK293_IR_first" = "#009ACD", "CCD841" = "#009ACD", "BJTERT" = "#009ACD",
                      "RMS444" = "#FAFA33", "RMS012" = "#FAFA33", "RMS007" = "#FFBF00","RMS006" = "#ff0000",
                      "RMS410" = "#ff0000","RMS127" = "#ff0000", "RMS000HQC" = "#ff0000",
                      "RMS102" = "#ff0000","RMS102_IR_first" = "#ff0000"))


# Add row and column annotations
row_ha = anno_mark(at = c(1:28), labels = df$Mechanism, which = "row")
column_ha = HeatmapAnnotation(foo = c(0,0.2,0.2,1,0.2,0.2,0.2,0.2,0.2,0.2,0.2,0.2,1), bar = colnames(df_m), col = colors  )

# Plot a heatmap
pdf("FIG_4B_ComplexHeatmap_dDSS_validation.pdf", width = 10, height = 8)
ht <- Heatmap(df_m,
              cluster_rows=FALSE,
              cluster_columns=FALSE,
              col=colorRamp2(c(-15, 0, 15),c("blue", "white", "red")),
              top_annotation = column_ha,
              heatmap_legend_param = list(
                legend_direction = "horizontal", 
                legend_width = unit(5, "cm")
              )
) + 
  rowAnnotation(mark = row_ha)

draw(ht, auto_adjust = FALSE,heatmap_legend_side = "bottom")
dev.off()
```

## Figure 4C
Gene centric validation of DSS correlation with BCL2L1 and PLK1 - All tumoroids
```{r, FIG 4C}
setwd()

# Choose data
df <- read.table("Validation_Paired_Wilcoxon_BCL_PLK1_xL_long.csv",sep=",", header = TRUE)

# Subset data into models and irradiation status
df_RMS <- subset(df,Sample == "RMS")
df_RMS_irr <- subset(df_RMS,Status == "Irradiated")
df_RMS_NonIrr <- subset(df_RMS,Status == "NonIrr")

# Perform a wilcoxon rank-sum test
result_RMS = wilcox.test(df_RMS_NonIrr$DSS, df_RMS_irr$DSS, paired = TRUE)

# State groups for comparison
my_comparisons <- rev(list(c("NonIrr","Irradiated")))

# Define data frame columns into correct types
df_RMS$Status <- factor(df_RMS$Status, levels = c("NonIrr","Irradiated"))

# Plot boxplots
ggplot(df_RMS, aes(x=Status, y=DSS, fill = Status)) + 
  geom_boxplot(outlier.shape = NA, position=position_dodge(1)) +
  scale_fill_brewer(palette="Dark2")+
  geom_jitter(color="black", size=2, width = 0.2) +
  stat_compare_means(comparisons = my_comparisons, method = "wilcox.test",paired=TRUE) +
  facet_wrap(~Gene,ncol=2) +
  theme(axis.line = element_line(colour = "black"),legend.position="none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background = element_rect(fill = 'white')) 

# Save plots
ggsave(filename = "FIG_4C_paired_Wilcoxon_NB_all.pdf", width = 4, height = 2, units = "in", device='pdf')
```

## Figure 4D
Gene centric validation of DSS correlation with BCL2L1 and PLK1 - Tumoroids in validation screen only
```{r, FIG 4D}
setwd()

# Choose data
df <- read.table("Validation_Paired_Wilcoxon_BCL_PLK1_xL_long.csv",sep=",", header = TRUE)

# Subset data into models and irradiation status
df_RMS <- subset(df,Sample == "RMS")
df_RMS_irr <- subset(df_RMS,Status == "Irradiated")
df_RMS_NonIrr <- subset(df_RMS,Status == "NonIrr")

df_RMS_validation <- subset(df_RMS,First_screened == "Validation")
df_RMS_validation_irr <- subset(df_RMS_validation,Status == "Irradiated")
df_RMS_validation_NonIrr <- subset(df_RMS_validation,Status == "NonIrr")

# Perform a wilcoxon rank-sum test
result_RMS_validation = wilcox.test(df_RMS_validation_NonIrr$DSS, df_RMS_validation_irr$DSS, paired = TRUE)

# State groups for comparison
my_comparisons <- rev(list(c("NonIrr","Irradiated")))

# Define data frame columns into correct types
df_RMS_validation$Status <- factor(df_RMS_validation$Status, levels = c("NonIrr","Irradiated"))

# Plot boxplots
ggplot(df_RMS_validation, aes(x=Status, y=DSS, fill = Status)) + 
  geom_boxplot(outlier.shape = NA, position=position_dodge(1)) +
  scale_fill_brewer(palette="Dark2")+
  geom_jitter(color="black", size=2, width = 0.2) +
  stat_compare_means(comparisons = my_comparisons, method = "wilcox.test",paired=TRUE) +
  facet_wrap(~Gene,ncol=2) +
  theme(axis.line = element_line(colour = "black"),legend.position="none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background = element_rect(fill = 'white')) 

# Save plots
ggsave(filename = "FIG_4D_paired_Wilcoxon_NB_validation.pdf", width = 4, height = 2, units = "in", device='pdf')
```

## Figure 4E
Gene centric validation of DSS correlation with BCL2L1 and PLK1 - All non-cancer models
```{r, FIG 4E}
setwd()

# Choose data
df <- read.table("Validation_Paired_Wilcoxon_BCL_PLK1_xL_long.csv",sep=",", header = TRUE)

# Subset data into models and irradiation status
df_control <- subset(df,Sample == "Control")
df_control_irr <- subset(df_control,Status == "Irradiated")
df_control_NonIrr <- subset(df_control,Status == "NonIrr")

# Perform a wilcoxon rank-sum test
result_control = wilcox.test(df_control_NonIrr$DSS, df_control_irr$DSS, paired = TRUE)

# State groups for comparison
my_comparisons <- rev(list(c("NonIrr","Irradiated")))

# Define data frame columns into correct types
df_control$Status <- factor(df_control$Status, levels = c("NonIrr","Irradiated"))

# Plot boxplots
ggplot(df_control, aes(x=Status, y=DSS, fill = Status)) + 
  geom_boxplot(outlier.shape = NA, position=position_dodge(1)) +
  scale_fill_brewer(palette="Dark2")+
  geom_jitter(color="black", size=2, width = 0.2) +
  stat_compare_means(comparisons = my_comparisons, method = "wilcox.test",paired=TRUE) +
  facet_wrap(~Gene,ncol=2) +
  theme(axis.line = element_line(colour = "black"),legend.position="none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background = element_rect(fill = 'white')) 

# Save plots
ggsave(filename = "FIG_4E_paired_Wilcoxon_control.pdf", width = 4, height = 2, units = "in", device='pdf')
```

## Figure 4F
DSS Comparison between PAX3 fusion positive and negative tumoroids - KI models
```{r, FIG 4F - Left}
setwd()

# Choose data
df <- read.csv("FIG_S4E_BCL_XL_validation.csv", header = TRUE)

# Rearrange dataset
long <- melt(df, id.vars = c("Sample", "Screen", "Type","Radiation","Translocation","PAX3","Use","RNA"))
long <- subset(long,Use == "Yes")

# Define group comparisons
my_comparisons <- rev(list(c("Yes","No")))

# Plot a boxplot 
ggplot(long, aes(x=PAX3, y=value)) + 
          geom_boxplot(outlier.shape = NA, position=position_dodge(1)) +
          geom_jitter(aes(color = variable), size = 0.8, alpha = 0.9, width = 0.1) +
          scale_color_manual(values = c(
                    "Navitoclax" = "#6a3d9a",
                    "A.1331852" = "#58b3db",  
                    "WEHI.539" = "#dbb358"    
          )) +
          ylim(-2, 50) +
          xlab("PAX3 Fusion") +
          ylab("DSS") + 
          ggtitle("KI Fusion") +
  stat_compare_means(comparisons = my_comparisons, method = "wilcox.test",paired=FALSE, label.y = 45) +
  theme(axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_rect(fill = 'white'),
        legend.position = "none") 

# Save plot
ggsave(filename = "FIG_4F_KI_Fusion_DSS_color.pdf", width = 2, height = 2, units = "in", device='pdf')
```

DSS Comparison between BCL2L1 high expression and low expression tumoroids - KI models
```{r, FIG 4F - Right}
setwd()

# Choose data
df <- read.csv("FIG_S4E_BCL_XL_validation.csv", header = TRUE)

# Rearrange and group dataset
long <- melt(df, id.vars = c("Sample", "Screen", "Type","Radiation","Translocation","PAX3","Use","RNA"))
long <- subset(long,Use == "Yes")
long$RNA <- recode(long$RNA, "LOW" = "Low", "HIGH" = "High")
long$RNA <- factor(long$RNA, levels = c("Low", "High"))

# Define group comparisons
my_comparisons <- rev(list(c("High","Low")))

# Plot a boxplot
ggplot(long, aes(x=RNA, y=value)) + 
          geom_boxplot(outlier.shape = NA, position=position_dodge(1)) +
          geom_jitter(aes(color = variable), size = 0.8, alpha = 0.9, width = 0.1) +
          scale_color_manual(values = c(
                    "Navitoclax" = "#6a3d9a",
                    "A.1331852" = "#58b3db",  
                    "WEHI.539" = "#dbb358"    
          )) +
          ylim(-2, 50) +
          xlab(expression(italic("BCL2L1") ~ "RNA Expression")) +
          ylab("DSS") + 
          ggtitle("KI RNA") +
  stat_compare_means(comparisons = my_comparisons, method = "wilcox.test",paired=FALSE, label.y = 45) +
  theme(axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_rect(fill = 'white'),
        legend.position = "none") 

# Save plot
ggsave(filename = "SP_Figures/FIG_4F_KI_RNA_DSS_color.pdf", width = 2, height = 2, units = "in", device='pdf')
```

## Figure 4G
DSS Comparison between PAX3 fusion positive and negative tumoroids - PMC Models
```{r, FIG 4G - Left}
setwd()

# Choose data
df <- read.csv("FIG_S4F_PMC_Navitoclax_fusionDSS.csv", header = TRUE)

# Select data from PMC
df <- subset(df,Sample == "PMC")

# Define groups for comparison
my_comparisons <- rev(list(c("Yes","No")))

# Plot a boxplot
ggplot(df, aes(x=PAX3_FP, y=Navitoclax)) + 
          geom_boxplot(outlier.shape = NA, position=position_dodge(1)) +
          geom_jitter(color = "#6a3d9a", size = 0.8, alpha = 0.9, width = 0.1) +
          ylim(-2, 50) +
          xlab("PAX3 Fusion") +
          ylab("DSS") +
          ggtitle("PMC Fusion") +
  stat_compare_means(comparisons = my_comparisons, method = "wilcox.test",paired=FALSE) + 
  theme(axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_rect(fill = 'white'))

# Save plot
ggsave(filename = "FIG_4G_PMC_Navitoclax_fusionDSS_color.pdf", width = 2, height = 2, units = "in", device='pdf')
```


DSS Comparison between BCL2L1 high expression and low expression tumoroids - PMC models
```{r, FIG 4G - Right}
# Choose data
df <- read.csv("FIG_S4F_PMC_Navitoclax_RNA.csv", header = TRUE)

# Subset data from PMC
df <- subset(df,Sample == "PMC")

# Rearrange and group data
df$RNA <- recode(df$RNA, "LOW" = "Low", "HIGH" = "High")
df$RNA <- factor(df$RNA, levels = c("Low", "High"))

# Define groups for comparisons
my_comparisons <- rev(list(c("High","Low")))

# Plot a boxplot
ggplot(df, aes(x=RNA, y=Navitoclax)) + 
          geom_boxplot(outlier.shape = NA, position=position_dodge(1)) +
          geom_jitter(color = "#6a3d9a", size = 0.8, alpha = 0.9, width = 0.1) +
          ylim(-2, 50) +
          xlab(expression(italic("BCL2L1") ~ "RNA Expression")) +
          ylab("DSS") +
          ggtitle("PMC RNA") +
  stat_compare_means(comparisons = my_comparisons, method = "wilcox.test",paired=FALSE) + 
  theme(axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_rect(fill = 'white')) 

# Save plot
ggsave(filename = "FIG_4G_PMC_Navitoclax_RNA_DSS_color.pdf", width = 2, height = 2, units = "in", device='pdf')
```

# Figure 5
## Figure 5A
Longterm radiosensitization heatmap
```{r, FIG 5A}
setwd()

# Choose the data
df <- read_excel("FIG_5A_Validation_longterm.xlsx")

# Extract only relevant dataframe
df_m <- as.matrix(df[, 6:16])

# Set the drugs to be the row names
rownames(df_m) <- df$DRUG_NAME

# Define colors for each model and treatment duration
colors = list(bar = c("RMS102_P" = "#cad59a",
                      "RMS102_V" = "#ff8886",
                      "RMS102_96" = "#d4b58c",
                      "RMS102_168" = "#b3966f",
                      "RMS006_P" = "#cad59a",
                      "RMS006_V" = "#ff8886",
                      "RMS006_96" = "#d4b58c", 
                      "RMS006_168" = "#b3966f",
                      "BJTERT_V" = "#ff8886",
                      "BJTERT_96" = "#d4b58c", 
                      "BJTERT_168" = "#b3966f"))

# Annotate the rows and columns
row_ha = anno_mark(at = c(1:4), labels = df$Mechanism, which = "row")
column_ha = HeatmapAnnotation(Screen = c("Tumoroids","Tumoroids","Tumoroids","Tumoroids","Tumoroids","Tumoroids","Tumoroids","Tumoroids", "Non-cancer", "Non-cancer", "Non-cancer"), bar = colnames(df_m), col = colors)

# Plot the heatmap
pdf("FIG_5A_Longterm_dDSS_validation.pdf", width = 8, height = 4)
ht <- Heatmap(df_m,
              cluster_rows=FALSE,
              cluster_columns=FALSE,
              col=colorRamp2(c(-35, 0, 35),c("blue", "white", "red")),
              top_annotation = column_ha,
              heatmap_legend_param = list(
                legend_direction = "horizontal", 
                legend_width = unit(5, "cm")
              )
) + 
  rowAnnotation(mark = row_ha)

draw(ht, auto_adjust = FALSE,heatmap_legend_side = "bottom")
dev.off()
```

## Figure 5C
Dose Response Curves for Longterm Sensitization - A-1331852
```{r, FIG 5C}
setwd()

# Choose the data
data <- "FIG_S5A_Validation_longterm_Curves.xlsx"

# Choose drug of Interest
drug <-"A-1331852"

# Choose timepoint for RMS102 - 96H
df_102_96 <- read_excel(data, sheet = "RMS102_96H")
df_102_96_EBRT <- read_excel(data, sheet = "RMS102_96H_IRR")
df_102_96_drug <- subset(df_102_96, DRUG_NAME == drug)
df_102_96_EBRT_drug <- subset(df_102_96_EBRT, DRUG_NAME == drug)
df_102_96_drug$Irradiated <- "no"
df_102_96_EBRT_drug$Irradiated <- "yes"
df_102_96_drug$RMS <- "RMS102"
df_102_96_EBRT_drug$RMS <- "RMS102"
df_102_96_drug$Time <- "96H"
df_102_96_EBRT_drug$Time <- "96H"
EC50_102_96 <- df_102_96_drug$EC50
EC50_102_96_EBRT <- df_102_96_EBRT_drug$EC50

# Choose timepoint for RMS102 - 168H
df_102_168 <- read_excel(data, sheet = "RMS102_168H")
df_102_168_EBRT <- read_excel(data, sheet = "RMS102_168H_IRR")
df_102_168_drug <- subset(df_102_168, DRUG_NAME == drug)
df_102_168_EBRT_drug <- subset(df_102_168_EBRT, DRUG_NAME == drug)
df_102_168_drug$Irradiated <- "no"
df_102_168_EBRT_drug$Irradiated <- "yes"
df_102_168_drug$RMS <- "RMS102"
df_102_168_EBRT_drug$RMS <- "RMS102"
df_102_168_drug$Time <- "168H"
df_102_168_EBRT_drug$Time <- "168H"
EC50_102_168 <- df_102_168_drug$EC50
EC50_102_168_EBRT <- df_102_168_EBRT_drug$EC50

# Choose timepoint for RMS006 - 96H
df_006_96 <- read_excel(data, sheet = "RMS006_96H")
df_006_96_EBRT <- read_excel(data, sheet = "RMS006_96H_IRR")
df_006_96_drug <- subset(df_006_96, DRUG_NAME == drug)
df_006_96_EBRT_drug <- subset(df_006_96_EBRT, DRUG_NAME == drug)
df_006_96_drug$Irradiated <- "no"
df_006_96_EBRT_drug$Irradiated <- "yes"
df_006_96_drug$RMS <- "RMS006"
df_006_96_EBRT_drug$RMS <- "RMS006"
df_006_96_drug$Time <- "96H"
df_006_96_EBRT_drug$Time <- "96H"
EC50_006_96 <- df_006_96_drug$EC50
EC50_006_96_EBRT <- df_006_96_EBRT_drug$EC50

# Choose timepoint for RMS006 - 168H
df_006_168 <- read_excel(data, sheet = "RMS006_168H")
df_006_168_EBRT <- read_excel(data, sheet = "RMS006_168H_IRR")
df_006_168_drug <- subset(df_006_168, DRUG_NAME == drug)
df_006_168_EBRT_drug <- subset(df_006_168_EBRT, DRUG_NAME == drug)
df_006_168_drug$Irradiated <- "no"
df_006_168_EBRT_drug$Irradiated <- "yes"
df_006_168_drug$RMS <- "RMS006"
df_006_168_EBRT_drug$RMS <- "RMS006"
df_006_168_drug$Time <- "168H"
df_006_168_EBRT_drug$Time <- "168H"
EC50_006_168 <- df_006_168_drug$EC50
EC50_006_168_EBRT <- df_006_168_EBRT_drug$EC50

# Choose timepoint for BJhTERT - 96H
df_BJTERT_96 <- read_excel(data, sheet = "BJTERT_96H")
df_BJTERT_96_EBRT <- read_excel(data, sheet = "BJTERT_96H_IRR")
df_BJTERT_96_drug <- subset(df_BJTERT_96, DRUG_NAME == drug)
df_BJTERT_96_EBRT_drug <- subset(df_BJTERT_96_EBRT, DRUG_NAME == drug)
df_BJTERT_96_drug$Irradiated <- "no"
df_BJTERT_96_EBRT_drug$Irradiated <- "yes"
df_BJTERT_96_drug$RMS <- "BJTERT"
df_BJTERT_96_EBRT_drug$RMS <- "BJTERT"
df_BJTERT_96_drug$Time <- "96H"
df_BJTERT_96_EBRT_drug$Time <- "96H"
EC50_BJTERT_96 <- df_BJTERT_96_drug$EC50
EC50_BJTERT_96_EBRT <- df_BJTERT_96_EBRT_drug$EC50

# Choose timepoint for BJhTERT - 168H
df_BJTERT_168 <- read_excel(data, sheet = "BJTERT_168H")
df_BJTERT_168_EBRT <- read_excel(data, sheet = "BJTERT_168H_IRR")
df_BJTERT_168_drug <- subset(df_BJTERT_168, DRUG_NAME == drug)
df_BJTERT_168_EBRT_drug <- subset(df_BJTERT_168_EBRT, DRUG_NAME == drug)
df_BJTERT_168_drug$Irradiated <- "no"
df_BJTERT_168_EBRT_drug$Irradiated <- "yes"
df_BJTERT_168_drug$RMS <- "BJTERT"
df_BJTERT_168_EBRT_drug$RMS <- "BJTERT"
df_BJTERT_168_drug$Time <- "168H"
df_BJTERT_168_EBRT_drug$Time <- "168H"
EC50_BJTERT_168 <- df_BJTERT_168_drug$EC50
EC50_BJTERT_168_EBRT <- df_BJTERT_168_EBRT_drug$EC50

# Choose a representative data frame
df_drug <- df_102_96_drug

# Combine all data
comb <- rbind(df_102_168_drug, 
              df_102_168_EBRT_drug, 
              df_102_96_drug, 
              df_102_96_EBRT_drug, 
              df_006_168_drug, 
              df_006_168_EBRT_drug, 
              df_006_96_drug, 
              df_006_96_EBRT_drug,
              df_BJTERT_168_drug,
              df_BJTERT_168_EBRT_drug,
              df_BJTERT_96_drug, 
              df_BJTERT_96_EBRT_drug)

# Specify the columns to keep
columns_to_keep <- c("ID", "DRUG_NAME", "ANALYSIS_NAME", "EC50", "SLOPE", 
                     "MAX", "MIN", "Min.Conc.tested", "Max.Conc.tested", 
                     "EC50_std_error","D1", "D2", "D3", "D4", "D5", "Irradiated", "RMS", "Time")

# Subset the data to keep only the specified columns
comb <- comb[, columns_to_keep]

# Reformat the data
long_comb <- melt(setDT(comb), id.vars = c("ID","DRUG_NAME","ANALYSIS_NAME","EC50","SLOPE","MAX","MIN",
                                           "Min.Conc.tested","Max.Conc.tested","EC50_std_error","Irradiated","RMS", "Time"), variable.name = "Viability")

# Add the concentration of the compound
long_comb$concentration <- c(
                             df_drug$Min.Conc.tested*1,
                             df_drug$Min.Conc.tested*1,
                             df_drug$Min.Conc.tested*1,
                             df_drug$Min.Conc.tested*1, 
                             df_drug$Min.Conc.tested*1,
                             df_drug$Min.Conc.tested*1,
                             df_drug$Min.Conc.tested*1,
                             df_drug$Min.Conc.tested*1,
                             df_drug$Min.Conc.tested*1,
                             df_drug$Min.Conc.tested*1,
                             df_drug$Min.Conc.tested*1,
                             df_drug$Min.Conc.tested*1,
                             df_drug$Min.Conc.tested*10,
                             df_drug$Min.Conc.tested*10,
                             df_drug$Min.Conc.tested*10,
                             df_drug$Min.Conc.tested*10,
                             df_drug$Min.Conc.tested*10,
                             df_drug$Min.Conc.tested*10,
                             df_drug$Min.Conc.tested*10,
                             df_drug$Min.Conc.tested*10,
                             df_drug$Min.Conc.tested*10,
                             df_drug$Min.Conc.tested*10,
                             df_drug$Min.Conc.tested*10,
                             df_drug$Min.Conc.tested*10,
                             df_drug$Min.Conc.tested*100,
                             df_drug$Min.Conc.tested*100,
                             df_drug$Min.Conc.tested*100,
                             df_drug$Min.Conc.tested*100,
                             df_drug$Min.Conc.tested*100,
                             df_drug$Min.Conc.tested*100,
                             df_drug$Min.Conc.tested*100,
                             df_drug$Min.Conc.tested*100,
                             df_drug$Min.Conc.tested*100,
                             df_drug$Min.Conc.tested*100,
                             df_drug$Min.Conc.tested*100,
                             df_drug$Min.Conc.tested*100,
                             df_drug$Min.Conc.tested*1000,
                             df_drug$Min.Conc.tested*1000,
                             df_drug$Min.Conc.tested*1000,
                             df_drug$Min.Conc.tested*1000,
                             df_drug$Min.Conc.tested*1000,
                             df_drug$Min.Conc.tested*1000,
                             df_drug$Min.Conc.tested*1000,
                             df_drug$Min.Conc.tested*1000,
                             df_drug$Min.Conc.tested*1000,
                             df_drug$Min.Conc.tested*1000,
                             df_drug$Min.Conc.tested*1000,
                             df_drug$Min.Conc.tested*1000,
                             df_drug$Min.Conc.tested*10000,
                             df_drug$Min.Conc.tested*10000,
                             df_drug$Min.Conc.tested*10000,
                             df_drug$Min.Conc.tested*10000,
                             df_drug$Min.Conc.tested*10000,
                             df_drug$Min.Conc.tested*10000,
                             df_drug$Min.Conc.tested*10000,
                             df_drug$Min.Conc.tested*10000,
                             df_drug$Min.Conc.tested*10000,
                             df_drug$Min.Conc.tested*10000,
                             df_drug$Min.Conc.tested*10000,
                             df_drug$Min.Conc.tested*10000)

# Ensure Time and Model is a factor with the desired order
long_comb$Time <- factor(long_comb$Time, levels = c("96H", "168H"))
long_comb$RMS <- factor(long_comb$RMS, levels = c("RMS006", "RMS102", "BJTERT"))

# Plot the dose reponse curves
ggplot(data = long_comb, aes(x = concentration, y = value, col = Irradiated)) + 
  geom_smooth(method = drm, method.args = list(fct = L.4()), se = FALSE, size = 1.5) +
  geom_point(size=1.5) +
  scale_color_manual(values=c("yes"="blue","no"="red"))+
  #ylim(-5, 190) +
  
  geom_vline(data=subset(long_comb, long_comb$RMS=="RMS006" & Time == "96H"), aes(xintercept=EC50_006_96),linetype="dotted", colour="red", linewidth = 1) + 
  geom_vline(data=subset(long_comb, long_comb$RMS=="RMS006" & Time == "96H"), aes(xintercept=EC50_006_96_EBRT),linetype="dotted", colour="blue", linewidth = 1) + 
  geom_vline(data=subset(long_comb, long_comb$RMS=="RMS006" & Time == "168H"), aes(xintercept=EC50_006_168),linetype="dotted", colour="red", linewidth = 1) + 
  geom_vline(data=subset(long_comb, long_comb$RMS=="RMS006" & Time == "168H"), aes(xintercept=EC50_006_168_EBRT),linetype="dotted", colour="blue", linewidth = 1) + 
  geom_vline(data=subset(long_comb, long_comb$RMS=="RMS102" & Time == "96H"), aes(xintercept=EC50_102_96),linetype="dotted", colour="red", linewidth = 1) + 
  geom_vline(data=subset(long_comb, long_comb$RMS=="RMS102" & Time == "96H"), aes(xintercept=EC50_102_96_EBRT),linetype="dotted", colour="blue", linewidth = 1) + 
  geom_vline(data=subset(long_comb, long_comb$RMS=="RMS102" & Time == "168H"), aes(xintercept=EC50_102_168),linetype="dotted", colour="red", linewidth = 1) + 
  geom_vline(data=subset(long_comb, long_comb$RMS=="RMS102" & Time == "168H"), aes(xintercept=EC50_102_168_EBRT),linetype="dotted", colour="blue", linewidth = 1) + 
  geom_vline(data=subset(long_comb, long_comb$RMS=="BJTERT" & Time == "96H"), aes(xintercept=EC50_BJTERT_96),linetype="dotted", colour="red", linewidth = 1) + 
  geom_vline(data=subset(long_comb, long_comb$RMS=="BJTERT" & Time == "96H"), aes(xintercept=EC50_BJTERT_96_EBRT),linetype="dotted", colour="blue", linewidth = 1) + 
  geom_vline(data=subset(long_comb, long_comb$RMS=="BJTERT" & Time == "168H"), aes(xintercept=EC50_BJTERT_168),linetype="dotted", colour="red", linewidth = 1) + 
  geom_vline(data=subset(long_comb, long_comb$RMS=="BJTERT" & Time == "168H"), aes(xintercept=EC50_BJTERT_168_EBRT),linetype="dotted", colour="blue", linewidth = 1) + 
  
 scale_x_log10(breaks = c(0.1, 1, 10, 100, 1000, 10000), labels = c("0.1", "1", "10", "100", "1000", "10000")) +
          
          theme(axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y = element_text(size = 10),
        axis.text.x = element_text(size = 10),
        axis.line.x = element_line(linewidth = 0.5),
        axis.line.y = element_line(linewidth = 0.5),
        panel.background = element_rect(fill = 'white'),
        legend.position = "none")+
          facet_grid(RMS ~ Time, scales = "free_y") +
          scale_y_continuous(limits = function(limits) {
                    c(-10, max(100, limits[2]))
          })

# Save the plot
ggsave(filename = paste0("FIG_S5A_Drug_curve_EBRT_",drug,"_vertical_EC50_ALL.pdf"), width = 4, height = 4, units = "in", device='pdf')

dev.off()
```

## Figure 5D
Survival fraction bar graph for colony formation assay
```{r, FIG 5D}
setwd()

# Choose the data
df <- read_excel("RMS_EBRT_CFA_R.xlsx")

# Define the sample
sample <- "RMS000HQC"

# Convert columns to correct types
df$Drug <- as.factor(df$Drug)
df$Gy <- as.numeric(as.character(df$Gy))

# Summarize mean and SD for each Gy and Drug
df_summary <- df %>%
  dplyr::mutate(Gy = factor(Gy)) %>%
  dplyr::group_by(Gy, Drug) %>%
  dplyr::summarise(
    Survival_mean = mean(Survival, na.rm = TRUE),
    Survival_sd = sd(Survival, na.rm = TRUE),
    .groups = "drop"
  )

# One-way ANOVA per Gy group
anova_results <- df %>%
  dplyr::group_by(Gy) %>%
  anova_test(Survival ~ Drug) %>%
  adjust_pvalue(method = "bonferroni")

# Filter only relevant comparisons against 0nM
posthoc <- df %>%
  group_by(Gy) %>%
  tukey_hsd(Survival ~ Drug) %>%
  filter(group1 == "0" | group2 == "0")  

# Custom vertical positions for significance symbols (avoid overlap)
posthoc_bar <- posthoc %>%
  mutate(
    comparison = paste0(group1, "_vs_", group2),
    y.position = case_when(
      Gy == "0" & comparison == "0_vs_1000" ~ 1.2,
      Gy == "0" & comparison == "0_vs_10"   ~ 1.1,
      Gy == "0" & comparison == "0_vs_0.1"  ~ 1.05,

      Gy == "2" & comparison == "0_vs_1000" ~ 0.25,
      Gy == "2" & comparison == "0_vs_10"   ~ 0.22,
      Gy == "2" & comparison == "0_vs_0.1"  ~ 0.20,

      Gy == "4" & comparison == "0_vs_1000" ~ 0.3,
      Gy == "4" & comparison == "0_vs_10"   ~ 0.2,
      Gy == "4" & comparison == "0_vs_0.1"  ~ 0.1,

      Gy == "6" & comparison == "0_vs_1000" ~ 0.3,
      Gy == "6" & comparison == "0_vs_10"   ~ 0.2,
      Gy == "6" & comparison == "0_vs_0.1"  ~ 0.1,

      TRUE ~ NA_real_
    )
  ) %>%
  filter(!is.na(y.position))  # Important: remove any rows where y.position is NA

#Plot the bar graph
ggplot(df_summary, aes(x = Drug, y = Survival_mean)) +
  geom_col(aes(fill = factor(Drug)), width = 0.6) +
  geom_errorbar(aes(ymin = Survival_mean - Survival_sd, ymax = Survival_mean + Survival_sd),
                width = 0.2) +
  facet_wrap(~ Gy, nrow = 1, labeller = as_labeller(c("0" = "0 Gy", "2" = "2 Gy", "4" = "4 Gy", "6" = "6 Gy"))) +
  stat_pvalue_manual(
    posthoc_bar,
    label = "p.adj.signif",
    tip.length = 0.01,
    hide.ns = TRUE
  ) +
  labs(
    title = paste0(sample),
    x = "A-1331852 (nM)",
    y = "Surviving Fraction"
  ) +
  scale_fill_manual(
    values = c(
      "0" = "#78DB95",     # green
      "0.1" = "#9C78DB",   # violet
      "10" = "#BAA5C9",    # warm lavender
      "1000" = "grey50"    # ash mauve
    )
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    legend.position = "none"
  )

# Save plot
ggsave(filename = paste0("FIG_5D_CFA_",sample,"_Bar.pdf"), width = 6, height = 3, units = "in", device='pdf')


```

## Figure 5E
Box plot for tumoroid growth assay and one-way ANOVA for endpoint (day 17)
```{r, FIG 5E}
setwd()

# Choose Data
df <- read_xlsx("SFA analysis_round2_FIJI.xlsx")

# Convert columns to correct types
df$Drug <- as.factor(df$Drug)
df$Gy <- as.factor(df$Gy)

# Filter only for 0, 0.1, and 1 nM
df_filtered <- df %>% filter(Drug %in% c("0", "0.1", "1"))

# One-way ANOVA per Gy
anova_results <- df_filtered %>%
  group_by(Gy) %>%
  anova_test(Fold ~ Drug)

# Post-hoc Tukey HSD per Gy
posthoc <- df_filtered %>%
  group_by(Gy) %>%
  tukey_hsd(Fold ~ Drug) %>%
  filter(group1 == "0") %>%   # Only compare against 0 nM
  add_y_position()

# Plot a boxplot
ggplot(df_filtered, aes(x = Drug, y = Fold)) +
  geom_boxplot(alpha = 0.8, outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 1.5, alpha = 0.6, aes(color = Drug)) +
  stat_pvalue_manual(
    posthoc,
    label = "p.adj.signif",
    tip.length = 0.01,
    hide.ns = TRUE
  ) +
  facet_wrap(~ Gy, labeller = as_labeller(gy_labels)) +  # consistent labelling
  scale_fill_manual(values = c(
    "1" = "#7895DB",  
    "0.1" = "#9C78DB",
    "0" = "#78DB95"
  )) +
  scale_color_manual(values = c(
    "1" = "#7895DB",  
    "0.1" = "#9C78DB",
    "0" = "#78DB95"
  )) +
  labs(
    x = "A-1331852 (nM)",
    y = "Fold Growth"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),                # remove gridlines
    axis.line = element_line(color = "black"),   # show axis lines
    axis.ticks = element_line(color = "black"),  # show axis ticks
    strip.text = element_text(size = 14),
    axis.text = element_text(size = 12),
    legend.position = "bottom",
    legend.title = element_blank()
  )

# Save plot
ggsave(filename = "FIG_5E_TGA_RMS006.pdf", width = 6, height = 4, units = "in", device='pdf')
```

